% -*- mode: latex; TeX-engine: xetex; LaTeX-command-style: (("" "SOURCE_DATE_EPOCH=0 %(PDF)%(latex) --shell-escape %S%(PDFout)")); TeX-master: "../dissertation.tex"; -*-

%% the abstract
Molecules, with their rich spectrum of internal states as well as
strong and tunable dipolar interactions, are promising candidates for
a broad range of experiments including quantum information,
quantum chemistry, precision measurement and probing physics beyond the Standard Model,
and quantum simulation of interacting many-body systems.
At the same time, the complexity that made them attractive also poses a significant challenge
to achieve a high level of quantum control that is required in many of
their potential applications.
Multiple approaches are being pursued for full quantum control of molecules
including association of atomic gases and direct cooling of molecules,
yet it remains challenging to control molecules at single particle level
due to the difficulty in tuning the atom-atom interaction
and laser cooling of molecules.

In this thesis, we achieved full control on a single weakly bound ground state molecule
created by coherent optical association from single atoms.
We first discuss the trapping and imaging of the atoms in the optical tweezer
and the cooling into a single quantum state in order to achieve the quantum control of atoms.
Combining this control with the resolution and flexibility of tweezers.
we developed a versatile platform to study interactions
between single atoms and properties of single molecules,
using precise optical spectroscopy on pairs of atoms in the tweezer.
Based on these studies, we proposed a promising pathway
for coherent all-optical creation of molecules
and experimentally demonstrated full control on the molecule including all degrees of freedom.
